Experimentally informed site-specific substitution models deepen phylogenetic estimates of the divergence of viral lineages

### 2018-02-06 manuscript/manuscript.tex
some tweaks to abstract

\begin{abstract}
\noindent \skhcomment{$\leq$ 250 words, currently 217}   
Molecular phylogenetics is often used to estimate the time since the divergence of modern gene sequences.
Such phylogenetic techniques often estimate substantially shallower divergence times than other methods.
For instance, in the case of viruses there is independent evidence that molecular phylogenetics can underestimate deep divergence times.  
This discrepancy is thought to be caused in part by inadequate models of purifying selection leading to branch-length underestimation.
Here we show that substitution models informed by experimental measurements of site-specific purifying selection lengthen the branches on phylogenies of influenza virus hemagglutinin.
This deepening of branch lengths is due to better modeling how site-specific amino-acid preferences affect the stationary state of the substitution models, and is independent of the branch-length-extension due to modeling site-to-site variation in substitution rate.
We show that the deepening of branch lengths from experimentally informed site-specific substitution models is similar to that achieved by other approaches that allow the stationary state to vary across sites.
However, the improvements from these site-specific models are limited by the inherent tension between the enhanced accuracy of accounting for site-specific amino-acid preferences and the fact that these preferences shift over long evolutionary times.
Overall, our work underscores the importance of modeling how site-specific purifying selection affects the stationary state as well as the substitution rate when estimating deep divergence times. 
\end{abstract}

>>>>

\begin{abstract}
\noindent  
Molecular phylogenetics is often used to estimate the time since the divergence of modern gene sequences.
Such phylogenetic techniques often estimate substantially shallower divergence times than other methods.
For instance, in the case of viruses there is independent evidence that molecular phylogenetics can underestimate deep divergence times.  
This discrepancy is thought to be caused in part by inadequate models of purifying selection leading to branch-length underestimation.
Here we show that substitution models informed by experimental measurements of the purifying selection due to site-specific amino-acid preferences lengthen the deep branches on phylogenies of influenza virus hemagglutinin.
This deepening of branch lengths is due to better modeling of the stationary state of the substitution models, and is independent of the branch-length-extension that results from modeling site-to-site variation in substitution rate.
The deepening of branch lengths from experimentally informed site-specific substitution models is similar to that achieved by other approaches that allow the stationary state to vary across sites.
However, the improvements from these site-specific models are limited by the inherent tension between the enhanced accuracy of accounting for site-specific amino-acid preferences and the fact that these preferences shift over long evolutionary times.
Overall, our work underscores the importance of modeling how site-specific purifying selection affects the stationary state when estimating deep divergence times. 
\end{abstract}
